---
title: "SPS 2016 Summer Bridge Data Science Math Final Project"
author: "Richard Pantoliano Jr"
date: "August 14, 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Overview
This project focuses on analyzing the hflights data set, which encompasses all flights that departed from Houston in 2011.

The two statistics this analysis will focus on are the amount of time each flight spent in the air, AirTime, and the arrival time of each flight, ArrTime. 

#Initialization
The Commands below load all of the libraries and data used in the analysis. The variable x will represent the 3rd quartile of the AirTime data, and the variable y will represent the mean of the ArrTime data.


```{r hflights x y}
#install.packages("hflights")
#install.packages("psych")
#install.packages("Hmisc")
#install.packages("MASS")
#install.packages("Lmoments")

require(psych)
require(hflights)
require(Hmisc)
require(MASS)
require(Lmoments)

data(hflights)

airsummary <- summary(hflights$AirTime)
arrsummary <-summary(hflights$ArrTime)

x <- as.integer(airsummary["3rd Qu."])
y <- as.integer(arrsummary["Mean"])

```

## Probability

a. The probability that the AirTime of a flight is > Q3 given that the ArrTime is greater than the mean is:
```{r}
#a. P(X>x | Y>y) = P(X>x & Y>y) / P(Y>y)
yg <- length(which(hflights$ArrTime > y)) / length(hflights$ArrTime)

xg_and_yg <- length(which(hflights$AirTime > x & hflights$ArrTime > y)) / length(hflights$AirTime)

pa <- xg_and_yg / yg
print(pa)
```

b. The probability that the AirTime of a flight is >  Q3 and that the ArrTime is greater than the mean is:
```{r}
#b. P(X>x, Y>y) = P(x>x & Y>y)
pb <- xg_and_yg
print(pb)
```

c. The probability that the AirTime of a flight is >  Q3 given that the ArrTime is greater than the mean is:
```{r}
#c. P(X<x | Y>y) = P(X<x & Y>y) / P(Y>y)
xl <- length(which(hflights$AirTime < x)) / length(hflights$AirTime)
xl_and_yg <- length(which(hflights$AirTime < x & hflights$ArrTime > y)) / length(hflights$AirTime)

pc <- xl_and_yg / yg
print(pc)
```

d. The probability that the AirTime of a flight is >  Q3 given that the ArrTime is < the mean is:
```{r}
#d. Since none was given, providing my own: P(X<x | Y<y) = P(X<x & Y<y) / P(Y<y)
yl <- length(which(hflights$ArrTime < y)) / length(hflights$ArrTime)
xl_and_yl <- length(which(hflights$AirTime < x & hflights$ArrTime < y)) / length(hflights$AirTime)

pd <- xl_and_yl / yl
print(pd)
```

Below, a matrix is constructed showing the counts of the listed events:
```{r}
#Count of x <= Q3 quartile and y <= Q2
xle3rd_yle2nd <- length( which(hflights$AirTime <= x & hflights$ArrTime <= y))
#print(xle3rd_yle2nd)

#Count of x > Q3 quartile and y <= Q2
xg3rd_yle2nd <- length( which(hflights$AirTime > x & hflights$ArrTime <= y))
#print(xg3rd_yle2nd)

#Count of x <= Q3 quartile and y > Q2
xle3rd_yg2nd <- length( which(hflights$AirTime <= x & hflights$ArrTime > y))
#print(xle3rd_yg2nd)

#Count of x > Q3 quartile and y > Q2
xg3rd_yg2nd <- length( which(hflights$AirTime > x & hflights$ArrTime > y))
#print(xg3rd_yg2nd)

t <- matrix(
    c(xle3rd_yle2nd,xg3rd_yle2nd, xle3rd_yle2nd + xg3rd_yle2nd,
      xle3rd_yg2nd, xg3rd_yg2nd, xle3rd_yg2nd + xg3rd_yg2nd,
      xle3rd_yle2nd + xle3rd_yg2nd, xg3rd_yle2nd + xg3rd_yg2nd,
      xle3rd_yle2nd + xg3rd_yle2nd+ xle3rd_yg2nd + xg3rd_yg2nd),
    ncol=3)
rownames(t) <- c("<=3d quartile", ">3d quartile", "total")
colnames(t) <- c("<=2d quartile", ">2d quartile", "total")
print(t)

```

This seems to show that the data is not independant.  Below,  let A be the new variable counting those observations above the 3d quartile for X, and let B be the new variable counting those observations for the 2d quartile for Y. The probability of P(A|B) if they are independent is given by prob_1.  The probability of P(A|B) if they are not independent is given by prob_2.

```{r}
# Test if P(A|B) = P(A)P(B) vs P(A|B) = P(A and B) / P(B)
prob_1 <- xl * yl # Probability if they are independant

# Probability if they are not independant
prob_2 <- length(which(hflights$AirTime < x & hflights$ArrTime < y))/  length(hflights$AirTime)

print(prob_1)
print(prob_2)
```

Using the Chi-squared test of independence:
```{r}
# Test for independance
print(chisq.test(hflights$AirTime, hflights$ArrTime))
```
This extremely small p-value indicates that the variables are dependant

# Descriptive and Inferential Statistics
Univariate info for AirTime:
```{r}
print(summary(hflights$AirTime))
print(describe(hflights$AirTime))

boxplot(hflights$AirTime,
        main = "Boxplot of Flight Air Time", 
        ylab="Total time in air",
        las=1
        )

```

Univariate info for ArrTime:
```{r}
print(summary(hflights$ArrTime))
print(describe(hflights$ArrTime))
boxplot(hflights$ArrTime,
        main = "Boxplot of Flight Arrival Time", 
        ylab="Arrival Time",
        las=1
)
```

Below is a scatterplot of the two variables
```{r}
plot(hflights$AirTime, hflights$ArrTime,
     main="Comparison of Time in Air vs Arrival Time",
     xlab="Time in Air",
     ylab="Arrival Time",
     las=1
     )
```

Below is the test with 95% CI of the difference in the mean of the two variables.  

```{r}
t.test(hflights$AirTime, hflights$ArrTime, paird=TRUE, conf.level=.95)
```

Below is the derived correlation matrix for the two variables.  
```{r}
data_subset <- subset(hflights, select=c("AirTime", "ArrTime"))
cor_matrix <- cor(data_subset, use="pairwise.complete.obs")
print(cor_matrix)
```

Below is the test with CI of 99% that the correlation between these variables is 0.
```{r}
cor.test(hflights$AirTime, hflights$ArrTime, conf.level=.99)
```
This indicates that the correlation is not zero.  That is, there is correlation between the variables.


# Linear Algebra
Inversion of the correlation matrix:
```{r}
inv_matrix <- solve(cor_matrix)
print(inv_matrix)
```

Below, both matrices are multiplied together in either order.
```{r}
print(inv_matrix%*%cor_matrix)
print(cor_matrix%*%inv_matrix)
```
Since the matricies are inverses, multiplying them in either order results in the identity matrix, which is the very definition of inverse matricies.

# Calculus-based Probability and statistics
We were asked to apply an exponential distribution to the skewed variable of the analyzed data.  In this case, we are focusing on the AirTime.

```{r}
airTime_clean <- na.omit(hflights$AirTime)
airrate = fitdistr(airTime_clean, densfun = "exponential")
print(airrate)
```

We can use the rate calculated from fitting the exponential function to the airTime data to randomly generate 1000 samples and create a histogram of that data.

```{r}
hist(rexp(1000,rate = airrate$estimate))
```

Below is the histogram of the original data.
```{r}
hist(airTime_clean)
```

As can be seen from the randomly generated data, the frequencies start with a peak at 0 and steadily decrease.  The recorded data increases up to the 150 minute mark, then decreases.  This stark difference indicates that the exponential function is a poor fit for modeling the observed data.