---
title: "SPS Week 2 R Assignment"
author: "Richard Pantoliano Jr"
date: "July 24, 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Implementing Factorial Recursively
Computes the value of n! recursively.
```{r recursive_factorial}
recursive_factorial <- function(input){
  if (input == 0) return(1)
  else return (input * recursive_factorial(input - 1))
}
```
Test examples:
```{r }
  recursive_factorial(0)
  recursive_factorial(1)
  recursive_factorial(2)
  recursive_factorial(3)
  recursive_factorial(4)
  recursive_factorial(15)

```


## Implementing Factorial Iteratively
Computes the value of n! iteratively.
```{r iterative_factorial}
iterative_factorial <- function(input){
  index <- input
  accumulator <- 1
  while (index > 0)
  {
    accumulator <- accumulator * index
    index <- index - 1
  }
  return(accumulator)
}
```
Test examples:
```{r }
  iterative_factorial(0)
  iterative_factorial(1)
  iterative_factorial(2)
  iterative_factorial(3)
  iterative_factorial(4)
  iterative_factorial(15)

```

## Using iterative_factorial and recursive_factorial to implement Combinations
The two functions defined below, iterative_choose and recursive_choose, make use of the iterative and recursive factorial functions to implement combinations.  Like the factorial functions, the choose functions should produce the same results

```{r iterative_choose}
iterative_choose <-function(elements, combination)
{
  return(iterative_factorial(elements) / (iterative_factorial(combination) * iterative_factorial(elements - combination) ) )
}
```

```{r recursive_choose}
recursive_choose<-function(elements, combination)
{
  return(recursive_factorial(elements) / (recursive_factorial(combination) * recursive_factorial(elements - combination) ) )
}
```

Test examples:
```{r }
  recursive_choose(5,2)
  iterative_choose(5,2)
  
  recursive_choose(10,1)
  iterative_choose(10,1)
  
  recursive_choose(3,3)
  iterative_choose(3,3)

```
