---
title: "Week 3 R Homework Assignment"
author: "Richard Pantoliano Jr"
date: "July 31, 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
## Transforming the Mushroom Dataset
We were assigned the task of transforming the CSV data located at https://archive.ics.uci.edu/ml/datasets/Mushroom using R.  Specifically, the code must:

1. pull the dataset from the Internet
2. transform the CSV data into a frame
3. create a subset of the data, including the "Class" column and 3-4 other columns
4. give meaningful names to the columns of the subset
5. replace the coded data in each column with the meaning contained within the dictionary for the dataset

## Pulling the dataset from the Internet
I used the RCurl library to pull the dataset from my personal Github repo.

```{r mushroom_data}
#install.packages("RCurl")
require(RCurl)
url <- "https://raw.githubusercontent.com/rpantoliano/CUNY_SPS_assignments/master/agaricus-lepiota.data"
mushroom_data <- getURL(url)
```

## Transform the CSV Data into a Data Frame
Following the download of the data, I used the read.table function to convert the CSV data into a data frame.  The parameters indicate that a text stream is being used as input instead of a file, that there is no header row, and that the separator for the data is a comma
```{r mushroom_df}
mushroom_df <- read.table(text=mushroom_data, header = FALSE, sep = ",")
head(mushroom_df)
```

## Create a Subset of the Data
I chose the columns V1, V3, V5, V11, and V17 from the dataset.  Using the subset function, I selected those columns by name to create a new data frame with only those columns.
```{r mushroom_ss}
mushroom_ss = subset(mushroom_df, select=c("V1","V3", "V5", "V11", "V17"))
head(mushroom_ss)
```

## Give the Columns Meaningful Names
Since there was no header to give names to the columns of the dataset, R gave the names V1:V23.  We will replace the names of the columns we selected with the names given in the data dictionary by using the rename function from the plyr library.
```{r}
#install.packages("plyr")
require(plyr)

mushroom_ss <- rename(mushroom_ss,
                      c("V1" = "Class",
                        "V3" = "Cap Surface",
                        "V5" = "Bruised",
                        "V11" = "Stalk Shape",
                        "V17" = "Veil Type"))
head(mushroom_ss)
```

## Replace the Coded Data
The CSV used single character codes for each type of data associated with each mushroom possibility.  Using the data dictionary included with the data, I used the sub function to replace each character code in each column with the associated descriptive information
```{r}
mushroom_ss$Class <- sub('^e$', "Edible", mushroom_ss$Class)
mushroom_ss$Class <- sub('^p$', "Poisonous", mushroom_ss$Class)

mushroom_ss$`Cap Surface` <- sub('^f$', "fibrous", mushroom_ss$'Cap Surface')
mushroom_ss$`Cap Surface` <- sub('^g$', "grooves", mushroom_ss$'Cap Surface')
mushroom_ss$`Cap Surface` <- sub('^y$', "scaly", mushroom_ss$'Cap Surface')
mushroom_ss$`Cap Surface` <- sub('^s$', "smooth", mushroom_ss$'Cap Surface')

mushroom_ss$Bruised <- sub('^t$', "bruised", mushroom_ss$Bruised)
mushroom_ss$Bruised <- sub('^f$', "no", mushroom_ss$Bruised)

mushroom_ss$`Stalk Shape` <- sub('^t$', "tapering", mushroom_ss$`Stalk Shape`)
mushroom_ss$`Stalk Shape` <- sub('^e$', "enlarging", mushroom_ss$`Stalk Shape`)

mushroom_ss$`Veil Type` <- sub('^p$', "partial", mushroom_ss$`Veil Type`)
mushroom_ss$`Veil Type` <- sub('^u$', "universal", mushroom_ss$`Veil Type`)

head(mushroom_ss, n = 20)
```
