---
title: "Week 4 R Homework Assignment"
author: "Richard Pantoliano Jr"
date: "August 7, 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Analyzing Electricity Rates
The following data was taken from the U.S. Federal Government's Open Data Website at <https://catalog.data.gov/dataset/u-s-electric-utility-companies-and-rates-look-up-by-zipcode-2013>.  Specifically, this data shows commercial, residential, and industrial electricity rates, by zip code across the USA.

The following code loads the necessary libraries into memory, pulls down a copy of the data from my own repository, and loads that data into a data frame.

```{r elec_df}
#install.packages("RCurl")
#install.packages("plyr")

require(RCurl)
require(plyr)

url <- "https://raw.githubusercontent.com/rpantoliano/CUNY_SPS_assignments/master/iouzipcodes2013.csv"

elec_data <- getURL(url)
elec_df <- read.csv(text=elec_data, header = TRUE, sep = ",")
```
## Comparing Commercial vs Residential rates
```{r echo=FALSE}
plot(elec_df$comm_rate, elec_df$res_rate, main="Comparison of Electrical Rates\n Commercial vs Residential", xlab="Commercial Rate", ylab="Residential Rate", las=1)
abline(lm(elec_df$comm_rate~elec_df$res_rate), col=4)

```

From this scatter plot, we can see that there is a positive correlation between the rates for commercial and residential electricity. The calculated correlation coefficient bares this out:
```{r}
cor(elec_df$comm_rate, elec_df$res_rate)
```


Below are the histograms for commercial and residential electrical rates:

```{r  echo = FALSE}
hist(elec_df$comm_rate,
     main = "Histogram of Commercial Electrical Rates",
     xlab = "Commercial Rate in kwh",
     ylab = "Frequency",
     las=1)
```

```{r echo = FALSE}
hist(elec_df$res_rate,
     main = "Histogram of Residential Electrical Rates",
     xlab = "Residential Rate in kwh",
     ylab = "Frequency",
     las=1)
```

These graphs indicate that residential rates are far more frequently higher than commercial rates.  That is, residential rates are mostly above $ .10 per kwh while most commercial rates are at or below $ .10 per kwh.


## Electrical Rates in NY
Below is a box plot of Commercial Electrical Rates in NY State.
```{r nydata, echo = FALSE}
nydata <- subset(elec_df, state=="NY")
boxplot(nydata$comm_rate,
        main = "Boxplot of Commercial Electrical Rates in NY", 
        xlab = "Commercial",
        ylab="Rate in kwh")
```

Below is a box plot of Residential Electrical Rates in NY State.
```{r echo = FALSE}
nydata <- subset(elec_df, state=="NY")
boxplot(nydata$res_rate,
        main = "Boxplot of Residential Electrical Rates in NY", 
        xlab = "Residential",
        ylab="Rate in kwh")

```

These two graphs show that the mean and max residential rates for electricity are higher than those rates for commercial electricity


