---
title: "Week 2 Data Math Homework"
author: "Richard Pantoliano Jr"
date: "July 24, 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Problem 1
P(pulling 505) = 1/540

## Problem 2
AEF, AEV, ATF, ATV,
CEF, CEV, CTF, CTV

## Problem 3
P(Heart) = 13/52
P(heart and !face) = 13/52 - 3/52 = 10/52 = 5/26

## Problem 4
P(sum < 6) = P(sum= 1) + P(sum= 2) + P(sum= 3) + P(sum= 4) + P(sum = 5)
          = 0 + 1/36 + 2/36 + 3/36 +  4/36
          = 10/36
          = 5/18
          
## Problem 5
```{r pmale}
males <- sum(233,159,102,220,250)
females <- sum(208,138,280,265,146)
pmale <- males / (females + males)
print(round(pmale, 4) )
```

## Problem 6
P(club, black card, face card) = 1/4 * 1/2 * 3/13 = 3/104

## Problem 7
P(spade | heart drawn first) = 13/51
        
## Problem 8
P(red | heart) = 12/51
        
## Problem 9
```{r}
total <- sum(12, 19, 12, 7, 12, 15, 4, 4)
print(( round(12 / (total - 1), 4 ) ) )
```

## Problem 10
Step 1: P(male and graduate degree) = 52 / 141

Step 2: P(male | graduate degree) = 52 / 102 = 26/51

## Problem 11
Number of Packages = 6 * 5 * 3 = 90

## Problem 12
Number of visits = 5! = 5 * 4 * 3 * 2 * 1 = 120

## Problem 13
Number of lineups = P(8,5) =
```{r perm}
perm <- function(n, k){choose(n, k ) * factorial(k ) }
print(perm(8, 5 ) )
```


## Problem 14
```{r}
print(choose(9,5) + choose(9,3) + choose(9,5) + choose(9,1))
```

## Problem 15
Number of 6 combinations of 14 toppings = C(14, 6)
```{r}
print(choose(14, 6 ) )
```

## Problem 16
Number of 3 card combinations drawn = C(52, 3)
```{r}
print(choose(52, 3 ) )
```

## Problem 17
Number of theater systems = 12 tvs * 9 sound systems * 5 dvd players = 540

## Problem 18
Number of passwords = P(26,5) * P(10,3)
```{r}
print(perm(26, 5 ) * perm(10, 3 ) )
```

## Problem 19
P(9,4)
```{r}
print(perm(9, 4 ) )
```

## Problem 20
C(11,8)
```{r}
print(choose(11, 8 ) )
```

## Problem 21
P(12,8) / C(12,4)
```{r}
print(perm( 12, 8) / choose(12, 4 ))
```

## Problem 22
P(13, 7)
```{r}
print( perm(13,7 ) )
```

## Problem 23
Since the problem capitalizes the first 'p' in "Population", I will treat them as distinct from one another. Given that, 
P(10,10) = 10!
```{r}
print(factorial(10 ) )
```

##Problem 24
Step 1: Find the expected value
```{r x px mu}
x <- c(5, 6, 7, 8, 9 )
px <- c(.1, .2, .3, .2, .2)
mu <- sum( x * px )
print(mu)
```

Step 2: Find the variance
```{r setvar}
setvar <- sum( (x - mu ) ^ 2 * px )
print(setvar)
```

Step 3: Find the standard deviation
```{r setstddev}
setstddev <- sqrt(setvar)
print(round(setstddev, 1 ) )
```

Step 4: Find value of P(X>=9)
P(X>=9) = .2

Step 5: Find P(X <= 7)
P(X <= 7) = .1 + .2 + .3 = .6

## Problem 25
Step 1: Find the expected value

```{r exptvalue}
exptvalue <- 23 * 188 / 376 - 4 * (376 - 188 ) / 376 
print(round(exptvalue, 2))

```

Step 2: Repeat experiment 994
```{r}
print(round(994 * exptvalue,2))

```

## Problem 26
Step 1:
```{r}
success <-pbinom(8,11,.5)
print(round(1 * success - 7 * (1-success), 2))
```
Step 2:
```{r}
print(round(632 *(1 * success - 7 * (1 - success ) ), 2))
```


## Problem 27
Step 1:
```{r exptvalue1}
exptvalue1 <- 583 * 12 / 51 - 35 * 39 / 51 
print(round(exptvalue1, 2))

```

Step 2:
```{r}
print(round(632 * exptvalue1,2))

```



## Problem 28
Modeling using the binomial distribution
```{r}
print(round(1 - pbinom(2, 10, .3 ), 3 ) )
```

## Problem 29
```{r}
print(5 * .3 )
```

## Problem 30
P(> 5) = 1 - P(=<5)
```{r}
print(round(1 - ppois(5, 5.5), 3))
```

## Problem 31
P(> 4) = 1 - P(<=3)
```{r}
print(round(1 - ppois(3, 5.7), 4))
```

## Problem 32
P(<=1)
```{r}
print(round(pbinom(1, 7, .4), 4))
```

## Problem 33
P(> 1) = 1 - P(<=1) = 1 - (P(0) + P(1))
```{r}
print( round(1 - ( perm(19, 8 ) + 19 * perm(19, 7 ) ) / perm (25, 8 ), 3))
```

## Problem 34
P(<7)
```{r}
print(round(pbinom(6, 8, .4), 3))
```







