---
title: "Week 4 Math Homework Assignment"
author: "Richard Pantoliano Jr"
date: "August 7, 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## 1. Using matrix operations, describe the solutions for the following family of equations:

x + 2y - 3z = 5

2x + y - 3z = 13

-x + y + 2z= -8

$$\begin{array}
{rrr}
1 & 2 & -3 & 5 \\
2 & 1 & -3 & 13 \\
-1 & 1 & 2 & -8
\end{array}
$$
Steps to solve:

* R1 = R1 + R3
* R2 = R2 + 2*R3
* R1 = R1 + R2
* R1 = 1/6 * R1
* R2 = R2 - 3 * R1
* R3 = R3 - R1
* R3 = R3 - 2 * R2
* R3 = -1 * R3
* Swap R1 and R3
* Swap R2 and R3

$$\begin{array}
{rrr}
1 & 0 & 0 & 7 \\
0 & 1 & 0 & -1 \\
0 & 0 & 1 & 0
\end{array}
$$


#a.  Find the inverse of the above 3x3 (non-augmented) matrix.
```{r x y}
x <- matrix(c(1,2,-1,2,1,1,-3,-3,2), 3,3)
print(x)
y <- solve(x)
print(y)
```

#b.  Solve for the solution using R.
```{r}
z <- matrix(c(5,13,-8), 3,1)
print(y %*% z)

```

#c.  Modify the  3x3 matrix such that there exists only one non-zero variable in the solution set.
$$\begin{array}
{rrr}
1 & 2 & -3 & 5 \\
0 & 1 & -3 & 13 \\
0 & 1 & 2 & -8
\end{array}
$$


## 2.  Consider the matrix, q=matrix(c(3,1,4,4,3,3,2,3,2),nrow=3). Let b=c(1,4,5).  Use Cramer's rule and R to determine the solution, x, to qx=b, if one exists.  Show all determinants.

Here is the solution using only the inverse matrix:
```{r q b}
q=matrix(c(3,1,4,4,3,3,2,3,2),nrow=3)
b=c(1,4,5)
print(solve(q)%*%b)
print(q)
```

Here, we attempt to solve using Cramer's Rule.  First, we create matrices, replacing the respective vectors with the solution matrix
```{r qx qy qz}
qx <- q
qy <- q
qz <- q
qx[,1] <- b
qy[,2] <- b
qz[,3] <- b
```

Next, we find the determinant of each of the matrices.
```{r dq, dx dy dz}
dq <- det(q)
dx <- det(qx)
dy <- det(qy)
dz <- det(qz)

print(dq)
print(dx)
print(dy)
print(dz)
```

Finally, we calculate the values of each value:
```{r}
x <- dx / dq
y <- dy / dq
z <- dz / dq

print(x)
print(y)
print(z)
```

As shown above, the solution using Cramer's Rule and the solution using the inverse matrix are equivalent






